### 2015-07-27 text/ms.Rmd
Incorporate more comments

\noindent
^2^National Center for Ecological Analysis and Synthesis

@@

We first developed and tested ensemble methods on a fully factorial simulated dataset [@rosenberg2014] with precisely known status. These published simulated data included the following scenarios: three fish life histories: small pelagic, demersal, and large pelagic; three levels of biomass depletion at the start of the dataset compared to carrying capacity: 100%, 70%, and 40% of carrying capacity; and four harvest dynamic settings: (1) a constant harvest rate, (2) a harvest rate coupled with biomass to mimic an open-access single-species fishery, (3) a scenario where harvest rate increases continuously, and (4) a 'roller-coaster' scenario where the harvest rate increases and then decreases. Process noise (i.e. variability in population dynamics) was introduced to the models through two levels of multiplicative recruitment variability: $N(0, 0.2)$ and $N(0, 0.6)$, and was either uncorrelated or had a first-order autoregressive correlation of $0.6$. The simulation also included a scenario with multiplicative observation error around catch at $N(0, 0.2)$ and one without observation error. 
<!--TODO: make last sentence clearer
4. First paragraph in Methods describing simulations line 113 in pdf - you may want to describe the recrutiment variability a bit more. I know its not an important detail, but because its in the first paragraph of methods, it might turn people off reading further. Most people wouldn't understand what 'multiplcative' (including me) is or may not know what 'process error/noise' refers to. 
5. First paragraph in Methods last sentence - needed a bit of rewording because as was, i got quite confused. hopefully this new sentence makes sense and doesnt change the meaning. 
-->

@@

We also tested our ensemble methods on the RAM Legacy Stock Assessment Database [@ricard2012]. Our analysis of the stock-assessment database was based on version XX downloaded on XX (TODO). After removing stocks for which the models described below did not converge (TODO: add number), this database included `r ram_stocks_n` stocks.
<!-- Dan: Do we address the implications of only running our method on stocks that converged later on? i.e. is there something about those omitted stocks that would make them less suited for this type of analysis?-->

@@

A critical component to any predictive modelling exercise is to evaluate the performance of a model on new data [@hastie2009]. We used repeated three-fold cross validation to test predictive performance: we randomly divided the dataset into three sets, built superensemble models on two-thirds of the data, and evaluated predictive performance on the remaining third. We repeated this across each of the three splits and then repeated the whole procedure 50 times to account for bias that may result from any one set of validation splits. In the simulated dataset, there were 10 replicates of each unique combination of simulation parameters that differed only in stochastic variability. Since the dynamics of these populations were often similar, we grouped these stocks in the cross-validation process into either the training or testing split. When applied to the RAM Legacy Stock Assessment database, we used superensembles trained on the entire simulation dataset. However, since mPRM is built on the same stock assessment database, we refit mPRM on each training split.

Predictive performance can be evaluated with metrics that represent a variety of modelling goals. For continuous response variables such as the mean and slope of population status, performance metrics often measure some form of bias, precision, accuracy (a combination of bias and precision), or the ability to correctly rank or correlate across populations [e.g. @walther2005]. Here, we measure proportional error, defined as $(\hat{b} - b)/b$, where $\hat{b}$ and $b$ represent estimated and 'true' (or stock-assessed) mean or slope of $B/B_\mathrm{MSY}$. We calculated median proportional error to measure bias, median absolute proportional error to measure accuracy, and Spearman's rank-order correlation between predicted and 'true' values to measure the ability to correctly rank populations. We measured the ability for each model to correctly place stocks into categories of above or below $B/B_\mathrm{MSY} = 1$ at any decision threshold by plotting receiver operating characteristic curves. Code to reproduce our analysis is available at <https://github.com/datalimited/ensembles> (TODO: make public). When cross-validating against the RAM Legacy Stock Assessment database, we treat the estimates from these data-rich stock assessments as known without error. Thus, any error in the stock assessment estimates of the mean or slope of $B/B_\mathrm{MSY}$ also contributes to our estimates of prediction error for each of the four data-limited models and the ensembles.

*TODO Olaf: B/Bmsy = 0.5 may be a better threshold as this corresponds to the overfished definition in the US and a few other places.  As Ray always points out, if Bmsy is the target, then you’d expect to be below it about half of the time in a well managed fishery.*

@@

Applied to the simulated dataset of known status, the individual models had variable success at recovering the mean (status) and slope (trajectory) of $B/B_\mathrm{MSY}$ in the last five years. All models exhibited a high degree of scatter around the one-to-one line of perfect status prediction (Fig. \ref{hexagon}). In contrast to the known unimodal distribution of status, CMSY exhibited bimodal predictions (Fig. \ref{hexagon}a), but had the best rank-order correlation and accuracy scores (Fig. \ref{performance}a). COM-SIR and SSCOM both correctly identified a number of stocks with low status, but frequently predicted a high status when status was in fact low (Fig. \ref{hexagon}b, c). mPRM had relatively poor ability to predict status for the simulated dataset (Fig. \ref{hexagon}d). There was generally little correlation between true and predicted trajectory for any of the individual models (rank-order correlation = `r range_slope_corr_non_sscom`) with the exception of SSCOM (correlation = `r slope_corr_sscom`; Figs \ref{scatter-sim-slope}a--d).

@@

When applied to the stock assessment database, the superensemble models --- trained exclusively on the simulated dataset --- generally performed as well or better than the best individual models. The mean, random forest, and GBM ensembles outperformed the mPRM method which is trained directly on the RAM Legacy Stock Assessment database itself (Fig. \ref{performance}b, Fig. \ref{hexagon-ram}). Compared to the individual models, the machine learning superensembles increased accuracy by `r mean_ram$mach_vs_ind_mare_fold`%, improved correlation from `r mean_ram$ind_corr_range` to `r mean_ram$mach_corr_range`, and reduced bias from `r mean_ram$ind_mre_range` to `r mean_ram$mach_mre_range`.

>>>>

\noindent
^2^National Center for Ecological Analysis and Synthesis, University of
California Santa Barbara, 735 State Street, Suite 300, Santa Barbara, CA 93103,
USA

@@

We first developed and tested ensemble methods on a fully factorial simulated dataset [@rosenberg2014] with precisely known status. These published simulated data included the following scenarios: three fish life histories: small pelagic, demersal, and large pelagic; three levels of biomass depletion at the start of the dataset compared to carrying capacity: 100%, 70%, and 40% of carrying capacity; and four harvest dynamic settings: (1) a constant harvest rate, (2) a harvest rate coupled with biomass to mimic an open-access single-species fishery, (3) a scenario where harvest rate increases continuously, and (4) a 'roller-coaster' scenario where the harvest rate increases and then decreases. Process noise (i.e. variability in population dynamics) was introduced to the models through two levels of multiplicative recruitment variability: $N(0, 0.2)$ and $N(0, 0.6)$, and was either uncorrelated or had a first-order autoregressive correlation of $0.6$. The simulation also included a scenario with multiplicative observation error around catch at $N(0, 0.2)$ and one without observation error. 
<!--TODO: make last sentence clearer
4. First paragraph in Methods describing simulations line 113 in pdf - you may want to describe the recrutiment variability a bit more. I know its not an important detail, but because its in the first paragraph of methods, it might turn people off reading further. Most people wouldn't understand what 'multiplcative' (including me) is or may not know what 'process error/noise' refers to. 
5. First paragraph in Methods last sentence - needed a bit of rewording because as was, i got quite confused. hopefully this new sentence makes sense and doesnt change the meaning. 
TODO: Jamie: add table instead / in addition to this paragraph?
-->

@@

We also tested our ensemble methods on the RAM Legacy Stock Assessment Database [@ricard2012] --- a compilation of stock-assessment output from hundreds of exploited marine populations around the world. Our analysis of the stock-assessment database was based on version XX downloaded on XX (TODO). After removing stocks for which the models described below did not converge (TODO: add number), this database included `r ram_stocks_n` stocks.
<!-- Dan: Do we address the implications of only running our method on stocks that converged later on? i.e. is there something about those omitted stocks that would make them less suited for this type of analysis?-->

@@

A critical component to any predictive modelling exercise is to evaluate the performance of a model on new data [@hastie2009]. We used repeated three-fold cross validation to test predictive performance: we randomly divided the dataset into three sets, built superensemble models on two-thirds of the data, and evaluated predictive performance on the remaining third. We repeated this across each of the three splits and then repeated the whole procedure 50 times to account for bias that may result from any one set of validation splits. In the simulated dataset, there were 10 replicates of each unique combination of simulation parameters that differed only in stochastic variability. Since the dynamics of these populations were often similar, we grouped these stocks in the cross-validation process into either the training or testing split. When applied to the RAM Legacy Stock Assessment database, we used superensembles trained on the entire simulation dataset. However, since mPRM is built on the same stock-assessment database, we refit mPRM on each training split.

Predictive performance can be evaluated with metrics that represent a variety of modelling goals. For continuous response variables such as the mean and slope of population status, performance metrics often measure some form of bias, precision, accuracy (a combination of bias and precision), or the ability to correctly rank or correlate across populations [e.g. @walther2005]. Here, we measure proportional error, defined as $(\hat{b} - b)/\abs{b}$, where $\hat{b}$ and $b$ represent estimated and 'true' (or stock-assessed) mean or slope of $B/B_\mathrm{MSY}$. We calculated median proportional error to measure bias, median absolute proportional error to measure accuracy, and Spearman's rank-order correlation between predicted and 'true' values to measure the ability to correctly rank populations. When cross-validating against the RAM Legacy Stock Assessment database, we treated the estimates from these data-rich stock assessments as known without error. Thus, any error in the stock-assessment estimates of the mean or slope of $B/B_\mathrm{MSY}$ also contributes to our estimates of prediction error for each of the four data-limited models and the ensembles. We measured the ability for each model to correctly place stocks into categories of above or below $B/B_\mathrm{MSY} = 0.5$ at any decision threshold by plotting receiver operating characteristic curves; a threshold of $B/B_\mathrm{MSY} = 0.5$ is used to classify a stock as overfished in Australia and the United States [@daff2007; @hilborn2010]. Code to reproduce our analysis is available at <https://github.com/datalimited/ensembles> (TODO: make public).

@@

Applied to the simulated dataset of known stock status, the individual models had variable success at recovering the mean (status) and slope (trajectory) of $B/B_\mathrm{MSY}$ in the last five years. All models exhibited a high degree of scatter around the one-to-one line of perfect status prediction (Fig. \ref{hexagon}). In contrast to the known unimodal distribution of status, CMSY exhibited bimodal predictions (Fig. \ref{hexagon}a), but had the best rank-order correlation and accuracy scores (Fig. \ref{performance}a). COM-SIR and SSCOM both correctly identified a number of stocks with low status, but frequently predicted a high status when status was in fact low (Fig. \ref{hexagon}b, c). mPRM had relatively poor ability to predict status for the simulated dataset (Fig. \ref{hexagon}d). There was generally little correlation between true and predicted trajectory for any of the individual models (rank-order correlation = `r range_slope_corr_non_sscom`) with the exception of SSCOM (correlation = `r slope_corr_sscom`; Figs \ref{scatter-sim-slope}a--d).

@@

When applied to the stock-assessment database, the superensemble models --- trained exclusively on the simulated dataset --- generally performed as well or better than the best individual models. The mean, random forest, and GBM ensembles outperformed the mPRM method which is trained directly on the RAM Legacy Stock Assessment database itself (Fig. \ref{performance}b, Fig. \ref{hexagon-ram}). Compared to the individual models, the machine learning superensembles increased accuracy by `r mean_ram$mach_vs_ind_mare_fold`%, improved correlation from `r mean_ram$ind_corr_range` to `r mean_ram$mach_corr_range`, and reduced bias from `r mean_ram$ind_mre_range` to `r mean_ram$mach_mre_range`.
